---
title: "DS_EXAM_COVID_CODE"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/DataScience Exam/US_covid")
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(lubridate)
library(tsibble)
library(tsibbledata) 
library(feasts)
library(fpp3)
library(fabletools)
library(fable)
library(distributional)
library(lme4)
library(tis)
library(prophet)
library(rstan)
library(tseries)
library(directlabels)
```


```{r}
# loading in COVID data for all of US
us_total <- read_csv("us_total21may.csv")
demographics_ny21may <- read_csv("demographics_ny21may.csv")
ny_covid <- read_csv("ny_covid21may.csv")
restrictions <- read_csv("restrictions.csv")

us_total <- us_total %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(index = date)

# transforming variables 
us_total <- us_total %>% 
  mutate(pos_log = log(cases)) %>%
  mutate(death_log = log(deaths)) %>%
  mutate(pos_diff = difference(cases)) %>%
  mutate(death_diff = difference(deaths))%>%
  mutate(death_diff_diff = difference(death_diff))%>%
  mutate(pos_diff_diff = difference(pos_diff))

us_total <- us_total %>%
  mutate(death_season = difference(death_diff, 7))%>%
  mutate(death_season_diff = difference(death_season))

# calculating lambda for cases 
lambda_cases <- us_total %>%
  features(cases, features = guerrero) %>%
  pull(lambda_guerrero)
# mutating box-cox transformed values for cases 
us_total <- us_total %>%
  mutate(box_cox_cases = box_cox(cases, lambda_cases))

# calculating lambda for deaths
lambda_deaths <- us_total %>%
  features(deaths, features = guerrero) %>%
  pull(lambda_guerrero)
# mutating box-cox transformed values for deaths 
us_total <- us_total %>%
  mutate(box_cox_deaths = box_cox(deaths, lambda_deaths))

# loading in the data with individual states
ny_covid <- read_csv("ny_covid21may.csv")
ny_covid <- ny_covid %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(key= state, index = date)

# loading in the data with restrictions
restrictions <- read_csv("restrictions.csv")
restrictions <- restrictions %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(index = date)
```

#autoplotting different variables of interest
```{r}
#autoplotting different variables of interest
autoplot(us_total, cases)+
  theme_classic()+
  scale_x_date(breaks = date_breaks("months"),labels = date_format("%b"))
autoplot(us_total, deaths)+
  theme_classic()
autoplot(us_total, pos_diff)+
    theme_classic()+
  scale_x_date(breaks = date_breaks("months"),labels = date_format("%b"))# cases increase per day 
autoplot(us_total, death_diff)+
    theme_classic()+
  scale_x_date(breaks = date_breaks("months"),labels = date_format("%b"))

# inspecting other transformed variables 
autoplot(us_total, death_diff) # deaths increase per day 
autoplot(us_total, pos_diff_diff) # doubles differenced cases
autoplot(us_total, death_diff_diff) # doubles differenced deaths
autoplot(us_total, pos_log)
autoplot(us_total, box_cox_cases)
autoplot(us_total, box_cox_deaths)
autoplot(us_total, difference(box_cox_cases))
autoplot(us_total, difference(box_cox_deaths))

#plotting for each state
autoplot(ny_covid, pos_pop)+
  geom_dl(aes(label = state), method = list(dl.combine("last.points")), cex = 0.5)+
  labs(title = "Normalised Positive cases - US state")+
  ylab("Positive cases")+
  theme_minimal()
autoplot(ny_covid, death_pop)+
  geom_dl(aes(label = state), method = list(dl.combine("last.points")), cex = 0.5)+
  labs(title = "Normalised deaths - US state")+
  ylab("Deaths")+
  theme_minimal()
```


#autocorrelation plts for cases and deaths 
```{r}
us_total %>%
  ACF(cases, lag_max = 100)%>%
  autoplot()
us_total %>%
  PACF(cases, lag_max = 100)%>%
  autoplot()
us_total %>%
  ACF(deaths, lag_max = 100)%>%
  autoplot()
us_total %>%
  PACF(deaths, max_lag = 100)%>%
  autoplot()
```


# STL decomposition and features of cases and deaths for all of US
```{r}
# stl decomposition for cases 
dcmp_cases <- us_total %>%
  model(stl = STL(cases))
components(dcmp_cases) %>% autoplot() +
  labs(title = "STL decomposition of positive cases")+
  theme_minimal()

# stl decomposition for deaths
dcmp_deaths <- us_total %>%
  model(stl = STL(deaths))
components(dcmp_deaths)
components(dcmp_deaths) %>% autoplot()+
  labs(title = "STL decomposition of deaths")+
  theme_minimal()

# cases stl features - trend/seasonal strength 
us_total %>%
  features(cases, feat_stl)
# deaths stl features - trend/seasonal strength 
us_total %>%
  features(deaths, feat_stl)
```


# Splitting US total into train/test split 
```{r}
# Train / test split - approx 80/20 
train <- us_total[0:387,]
test <- us_total[388:483,]

#Making test and train set into stibbles 
train <- train %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(index = date)

test <- test %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(index = date)
```

# Manually Searching and evaluating ETS models for cases
```{r}
train_fit_cases <- train %>%
  model(
    `Holt's method` = ETS(cases ~ error("A") + trend("A") + season("N")),
    `Additive Holt-Winter's method` = ETS(cases ~ error("A") + trend("A") + season("A")),
    `Damped Holt's method` = ETS(cases ~ error("A") + trend("Ad") + season("N")),
    `Additive Damped Holt-Winter's method` = ETS(cases ~ error("A") + trend("Ad") + season("A")), 
    `Simple Exponential method` = ETS(cases ~ error("A") + trend("N") + season("N"))
    )
train_fit_cases %>%
  report()
glance(train_fit_cases) %>% arrange(AIC) %>% select(.model:BIC)
train_fit_cases %>%
  accuracy() %>%
  arrange(MAPE)
train_fc_cases <- train_fit_cases %>%
  forecast(h = 96)
train_fc_cases %>%
  autoplot(us_total, level = NULL) +
  labs(
    y = "Positive Cases",
    title = "Positive COVID 19 Cases in the US"
  ) +
  guides(colour = guide_legend(title = "Forecast"))+
  theme_minimal()
accuracy(train_fc_cases, test)%>%
  arrange(RMSE)
train_fit_cases %>%
  select(`Additive Holt-Winter's method`) %>%
  gg_tsresiduals()
```

# Manually searching for and evaluating best ETS model deaths
```{r}
train_fit_deaths <- train %>%
  model(
    `Holt's method` = ETS(deaths ~ error("A") + trend("A") + season("N")),
    `Additive Holt-Winter's method` = ETS(deaths ~ error("A") + trend("A") + season("A")),
    `Damped Holt's method` = ETS(deaths ~ error("A") + trend("Ad") + season("N")), 
    `Additive Damped Holt-Winter's method` = ETS(deaths ~ error("A") + trend("Ad") + season("A")), 
    `Simple Exponential method` = ETS(deaths ~ error("A") + trend("N") + season("N")))

train_fit_deaths %>%
  report()
train_fit_deaths %>%
  accuracy() %>%
  arrange(MAPE)
glance(train_fit_deaths) %>% arrange(AIC) %>% select(.model:BIC)

train_fc_deaths <- train_fit_deaths %>%
  forecast(h = 96)

train_fc_deaths %>%
  autoplot(us_total, level = NULL) +
  labs(
    y = "Nr of deaths from COVID 19",
    title = "Deaths from COVID 19 in the US"
  ) +
  guides(colour = guide_legend(title = "Forecast"))+
  theme_minimal()

accuracy(train_fc_deaths, test)%>%
  arrange(MAPE)

train_fit_deaths %>%
  select(`Additive Damped Holt-Winter's method`) %>%
  gg_tsresiduals()
```


# Manual/AUtomatic Search for best arima models for cases
```{r}
caf_fit_cases <- train %>%
  model(arima021 = ARIMA(cases ~ pdq(0,2,1) + PDQ(0,0,2)),
        arima111 = ARIMA(cases ~ pdq(1,1,1) + PDQ(2,1,0)),
        arima322 = ARIMA(cases ~ pdq(3,2,2) + PDQ(0,0,1)),
        stepwise = ARIMA(cases),
        search = ARIMA(cases, stepwise=FALSE, greedy=FALSE, approximation = FALSE))

caf_fit_cases %>% pivot_longer(everything(), names_to = "Model name",
                         values_to = "Orders")
glance(caf_fit_cases) %>% arrange(AICc) %>% select(.model:BIC)
caf_fit_cases %>%
  select(arima111) %>%
  gg_tsresiduals()
caf_fit_cases %>%
  augment()
caf_fit_cases %>%
  accuracy() %>%
  arrange(MASE)
fc_cases_arima <- caf_fit_cases %>%
  forecast(h = 96)
fc_cases_arima %>%
  autoplot(us_total)+
  theme_minimal()
fc_cases_arima %>% 
  accuracy(test)
fc_cases_arima %>%
  autoplot(us_total, level = NULL) +
  guides(colour = guide_legend(title = "Forecast"))+
  labs(y = "Positive Cases",title = "Arima Forecast of positive cases") +
  theme_minimal()
```
# Manual/automatic Search for best arima models for death
```{r}
caf_fit_deaths <- train %>%
  model(arima021 = ARIMA(deaths ~ pdq(0,2,1) + PDQ(0,0,2)),
        arima112 = ARIMA(deaths ~ pdq(1,1,2) + PDQ(0,1,1)),
        arima211 = ARIMA(deaths ~ pdq(2,1,1) + PDQ(1,1,2)),
        arima312 = ARIMA(deaths ~ pdq(3,1,2) + PDQ(0,1,1)),
        arima111 = ARIMA(deaths ~ pdq(1,1,1) + PDQ(2,1,0)),
        stepwise = ARIMA(deaths),
        search = ARIMA(deaths, stepwise=FALSE, approximation = FALSE))

caf_fit_deaths %>% pivot_longer(everything(), names_to = "Model name",
                         values_to = "Orders")

glance(caf_fit_deaths) %>% arrange(AICc) %>% select(.model:BIC)

fc_deaths_arima <- caf_fit_deaths %>%
  forecast(h = 96)
fc_deaths_arima %>%
  autoplot(us_total)+
  theme_minimal()
fc_deaths_arima %>% 
  accuracy(test)
fc_deaths_arima %>%
  autoplot(us_total, level = NULL) +
  guides(colour = guide_legend(title = "Forecast"))+
    labs(y = "Deaths",title = "Arima Forecast of Deaths") +
  theme_minimal()

caf_fit_deaths %>%
  select(search) %>%
  gg_tsresiduals()
```

# testing Manually selected best Arima and ETS models on cases
```{r}
fit <- train %>%
  model(
    arima111 = ARIMA(cases ~ pdq(1,1,1) + PDQ(2,1,0)),
    `Additive Holt-Winter's method` = ETS(cases ~ error("A") + trend("A") + season("A")))
fit
fit %>%
  select(arima111, `Additive Holt-Winter's method`) %>%
  coef()

fit %>%
  glance()
fit %>%
  select(arima111,`Additive Holt-Winter's method`) %>%
  report()
fit %>%
  augment()

fit %>%
  accuracy() %>%
  arrange(MASE)
fc <- fit %>%
  forecast(h = 96)
fc %>%
  autoplot(us_total)+
  theme_minimal()+
  labs(y = "Positive Cases",title = "Comparing Arima and ETS Forecast of cases") 
fc %>% 
  accuracy(test)%>% arrange(RMSE)

#for cases it seems that the arima model does best on the training data, however, the ets model does better on forecasting the test data 
# Mean absolute percentage error: MAPE
#ets: 0.47
#arima: 2.59

fit %>%
  select(`Additive Holt-Winter's method`) %>%
  gg_tsresiduals()+
  labs(title = "`Additive Holt-Winter's method` residuals for cases") 

fit %>%
  select(arima111) %>%
  gg_tsresiduals()+
  labs(title = "Arima(1,1,1)(2,1,0)` residuals for cases")  

augment(fit) %>%
  features(.resid, ljung_box, lag = 7, dof = 0)
```



# camparing manually selected best Arima and ETS models on deaths
```{r}
fit <- train %>%
  model(
    `Additive Holt-Winter's method` = ETS(deaths ~ error("A") + trend("A") + season("A")),  
    arima312 = ARIMA(deaths ~ pdq(3,1,2) + PDQ(0,1,1)))
fit %>%
  select(arima312,`Additive Holt-Winter's method`) %>%
  coef()
fit %>%
  glance()
fit %>%
  select(arima312,`Additive Holt-Winter's method`) %>%
  report()
fit %>%
  augment()
fit %>%
  accuracy() %>%
  arrange(RMSE)
fc <- fit %>%
  forecast(h = 96)
fc %>%
  autoplot(us_total)+
  theme_minimal()+
  labs(y = "Deaths",title = "Comparing Arima and ETS Forecast of deaths") 
fc %>%
  accuracy(test)%>%
  arrange(RMSE)

# for deaths it also seems that the ets model performs best on the test data 
# MAPE: ets 2.24
# MAPE: arima 4.79

fit %>%
  select(`Additive Holt-Winter's method`) %>%
  gg_tsresiduals()+
  labs(title = "`Additive Holt-Winter's method` residuals for deaths") 

fit %>%
  select(arima312) %>%
  gg_tsresiduals()+
  labs(title = "`Arima(3,1,2)(0,1,1)` residuals for deaths") 

augment(fit) %>%
  features(.resid, ljung_box, lag = 7, dof = 0)
```


# Camparing forecasting with automatically chosen best Arima and ETS models on cases
```{r}
fit_cases <- train %>%
  model(
    ets = ETS(cases),
    arima = ARIMA(cases, stepwise=FALSE, approximation = FALSE))
fit_cases %>%
  select(arima, ets) %>%
  coef()
fit_cases %>%
  glance()
fit_cases %>%
  select(arima, ets) %>%
  report()
fit_cases %>%
  augment()
fit_cases %>%
  accuracy() %>%
  arrange(MASE)
fc_cases <- fit_cases %>%
  forecast(h = 96)
fc_cases %>%
  autoplot(us_total)
fc_cases %>% 
  accuracy(test)

fit_cases %>%
  select(ets) %>%
  gg_tsresiduals()

fit_cases %>%
  select(arima) %>%
  gg_tsresiduals()
```

```{r}
fit_deaths <- train %>%
  model(
    ets = ETS(deaths),
    arima = ARIMA(deaths, stepwise=FALSE, approximation = FALSE))
fit_deaths %>%
  select(arima, ets) %>%
  coef()
fit_deaths %>%
  glance()
fit_deaths %>%
  select(arima, ets) %>%
  report()
fit_deaths %>%
  augment()
fit_deaths %>%
  accuracy() %>%
  arrange(MASE)
fc_deaths <- fit_deaths %>%
  forecast(h = 96)
fc_deaths %>%
  autoplot(us_total)
fc_deaths %>% 
  accuracy(test)

fit_deaths %>%
  select(ets) %>%
  gg_tsresiduals()

fit_deaths %>%
  select(arima) %>%
  gg_tsresiduals()
```





###------- SUBSETS OF State of Interest (SOI) ---------####
```{r}
STATES_OI <- subset(ny_covid, state=="NY" | state=="VT" | state=="ND")
# removing dates that are not shared by all three series 
NY <- subset(ny_covid, state=="NY")
NY <- subset(NY, date >= "2020-03-11")
ND <- subset(ny_covid, state=="ND")
ND <- subset(ND, date >= "2020-03-11")
VT <- subset(ny_covid, state=="VT")
VT <- subset(VT, date >= "2020-03-11")

STATES_OI <- rbind(NY, ND, VT)
STATES_OI <- STATES_OI %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(key=state, index = date)

#changing the cases variables to the variables normalised by population for better comparison 
STATES_OI$cases_raw <- STATES_OI$cases
STATES_OI$cases <- STATES_OI$pos_pop
STATES_OI$deaths_raw <- STATES_OI$deaths
STATES_OI$deaths <- STATES_OI$death_pop

STATES_OI <- STATES_OI %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(key=state, index = date)

# Plotting the positive cases  and death trajectories 
autoplot(STATES_OI, cases)+
  geom_dl(aes(label = state), method = list(dl.combine("last.points")), cex = 0.1)+
  theme_minimal()+
  labs(title = "Positive Cases per State")+
  ylab("Cases/Population Size")
autoplot(STATES_OI, deaths)+
  geom_dl(aes(label = state), method = list(dl.combine("last.points")), cex = 0.1)+
  theme_minimal()+
  labs(title = "Deaths per State")+
  ylab("Deaths/Population Size")
```
# STL and features for each state 
```{r}
# STL decomposition of the cases per state 
dcmp_SOI <- STATES_OI %>%
  model(stl = STL(cases))
components(dcmp_SOI)%>%
  features()
components(dcmp_SOI) %>% autoplot()+
  theme_minimal()
# STL decomposition of the deaths per state 
dcmp_SOI <- STATES_OI %>%
  model(stl = STL(deaths))
components(dcmp_SOI)
components(dcmp_SOI) %>% autoplot()+
  theme_minimal()

STATES_OI %>%
  features(cases, feat_stl)

STATES_OI %>%
  features(deaths, feat_stl)
```


# Train/Test split of states of interest 
```{r}
train_SOI <- STATES_OI %>%
  group_by(state) %>%
  slice(1:348)

test_SOI <- STATES_OI %>%
  group_by(state) %>%
  slice(349:435)

train_SOI <- train_SOI %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(key=state, index = date)

test_SOI <- test_SOI %>%
  mutate(date = as_date(date)) %>%
  as_tsibble(key=state, index = date)
```

# evaluating best ETS model for each state
```{r}
train_SOI_fit <- train_SOI %>%
  model(
    `Simple Exponential method` = ETS(cases ~ error("A") + trend("N") + season("N")),
    `Holt's method` = ETS(cases ~ error("A") + trend("A") + season("N")),
    `Additive Holt-Winter's method` = ETS(cases ~ error("A") + trend("A") + season("A")),
    `Damped Holt's method` = ETS(cases ~ error("A") + trend("Ad") + season("N")),
    `Additive Damped Holt-Winter's method` = ETS(cases ~ error("A") + trend("Ad") + season("A"))
    )

glance(train_SOI_fit) %>% arrange(AIC) %>% select(.model:BIC)
train_SOI_fit %>%
  accuracy() %>%
  arrange(RMSE)

train_SOI_fc <- train_SOI_fit %>%
  forecast(h = 86)

train_SOI_fc %>%
  autoplot(STATES_OI, level = NULL) +
  labs(
    y = "Positive Cases",
    title = "Positive COVID 19 Cases per state"
  ) +
  guides(colour = guide_legend(title = "Forecast"))+
  theme_minimal()

accuracy(train_SOI_fc, test_SOI) %>% arrange(MAPE)
```
# fitting and chosing the best ets model for deaths for each state 
```{r}
train_SOI_fit_deaths <- train_SOI %>%
  model(
    `Simple Exponential method` = ETS(deaths ~ error("A") + trend("N") + season("N")),
    `Holt's method` = ETS(deaths ~ error("A") + trend("A") + season("N")),
    `Additive Holt-Winter's method` = ETS(deaths ~ error("A") + trend("A") + season("A")),
    `Damped Holt's method` = ETS(deaths ~ error("A") + trend("Ad") + season("N")),
    `Additive Damped Holt-Winter's method` = ETS(deaths ~ error("A") + trend("Ad") + season("A"))
    )

glance(train_SOI_fit_deaths) %>% arrange(AIC) %>% select(.model:BIC)
train_SOI_fit_deaths %>%
  accuracy() %>%
  arrange(RMSE)

train_SOI_fc_deaths <- train_SOI_fit_deaths %>%
  forecast(h = 86)

train_SOI_fc_deaths %>%
  autoplot(STATES_OI, level = NULL) +
  labs(
    y = "Deaths/Population size",
    title = "COVID 19 Deaths per state forecast"
  ) +
  guides(colour = guide_legend(title = "Forecast"))+
  theme_minimal()

test_scores <- accuracy(train_SOI_fc_deaths, test_SOI) %>% arrange(MAPE)
```
# automatically fitting and chosing the best ets and arima model for cases for each state 
```{r}
fit_ARIMA_states_cases <- train_SOI %>%
  model(
    ets = ETS(cases),
    arima = ARIMA(cases, stepwise = FALSE, approximation = FALSE))

fit_ARIMA_states_cases

fit_ARIMA_states_cases %>%
  select(arima, ets) %>%
  coef()

fit_ARIMA_states_cases %>%
  glance()
fit_ARIMA_states_cases %>%
  select(arima, ets) %>%
  report()
fit_ARIMA_states_cases %>%
  augment()
augment
fit_ARIMA_states_cases %>%
  accuracy() %>%
  arrange(MASE)
fc_ARIMA_states_cases <- fit_ARIMA_states_cases %>%
  forecast(h = 86)
fc_ARIMA_states_cases %>%
  autoplot(STATES_OI)+
  theme_minimal()+
  labs(y = "Cases",title = "Comparing Arima and ETS Forecast of cases per each state") 

test_scores <- fc_ARIMA_states_cases %>% 
  accuracy(test_SOI)%>%
  arrange(MAPE)
```


# automatically fitting and chosing the best ets and arima model for deaths for each state 
```{r}

fit_ARIMA_states_deaths <- train_SOI %>%
  model(
    ets = ETS(deaths),
    arima = ARIMA(deaths, stepwise = FALSE, approximation = FALSE))
fit_ARIMA_states_cases

fit_ARIMA_states_deaths %>%
  select(arima, ets) %>%
  coef()

fit_ARIMA_states_deaths %>%
  glance()
fit_ARIMA_states_deaths %>%
  select(arima, ets) %>%
  report()
fit_ARIMA_states_deaths %>%
  augment()
fit_ARIMA_states_deaths %>%
  accuracy() %>%
  arrange(MASE)
fc_ARIMA_states_deaths <- fit_ARIMA_states_deaths %>%
  forecast(h = 86)
fc_ARIMA_states_deaths %>%
  autoplot(STATES_OI)+
  theme_minimal()+
  labs(y = "Deaths",title = "Comparing Arima and ETS Forecast of deaths per each state") 

test_scores2 <- fc_ARIMA_states_deaths %>% 
  accuracy(test_SOI)%>%
  arrange(MAPE)
```

```{r}
test_scores <- rbind(test_scores, test_scores2)

test_score %>% 
  filter(.model=="arima")%>%
  summarise(sum = sum(MAPE), mean=mean(MAPE))
#12.02825	4.009415	
test_score %>% 
  filter(.model=="ets")%>%
  summarise(sum = sum(MAPE), mean = mean(MAPE))
#10.3067	3.435566	

# The best performing type of forecasting model across the different series is, according to the mean MAPE and RMSE score, the ETS
#of the ETS and ARIMA model acoording to the mean MAPE and RMSE scores across the different time series, the model performing best is the ETS model 

test_score %>% 
  filter(.model=="arima")%>%
  summarise(sum = sum(RMSE), mean=mean(RMSE))
#0.1800321	0.0600107	
test_score %>% 
  filter(.model=="ets")%>%
  summarise(sum = sum(RMSE), mean=mean(RMSE))
#0.1505088	0.0501696	
```
# CASES - checking residuals for seperate state and model forecasts 
```{r}
STATES_OI %>%
  filter(state=="ND")%>%
  model(ETS(cases)) %>%
  gg_tsresiduals()+
    labs(title="ND - cases - ETS")
STATES_OI %>%
  filter(state=="ND")%>%
  model(ARIMA(cases)) %>%
  gg_tsresiduals()+
  labs(title="ND - cases - ARIMA")

STATES_OI %>%
  filter(state=="NY")%>%
  model(ETS(cases)) %>%
  gg_tsresiduals()+
    labs(title="NY - cases - ETS")
STATES_OI %>%
  filter(state=="NY")%>%
  model(ARIMA(cases)) %>%
  gg_tsresiduals()+
  labs(title="NY - cases - ARIMA")

STATES_OI %>%
  filter(state=="VT")%>%
  model(ETS(cases)) %>%
  gg_tsresiduals()+
    labs(title="VT - cases - ETS")
STATES_OI %>%
  filter(state=="VT")%>%
  model(ARIMA(cases)) %>%
  gg_tsresiduals()+
  labs(title="VT - cases - ARIMA")
```
# DEATHS - checking residuals for seperate state and model forecasts 
```{r}
STATES_OI %>%
  filter(state=="ND")%>%
  model(ETS(deaths)) %>%
  gg_tsresiduals()+
    labs(title="ND - deaths - ETS")
STATES_OI %>%
  filter(state=="ND")%>%
  model(ARIMA(deaths)) %>%
  gg_tsresiduals()+
  labs(title="ND - deaths - ARIMA")

STATES_OI %>%
  filter(state=="NY")%>%
  model(ETS(deaths)) %>%
  gg_tsresiduals()+
    labs(title="NY - deaths - ETS")
STATES_OI %>%
  filter(state=="NY")%>%
  model(ARIMA(deaths)) %>%
  gg_tsresiduals()+
  labs(title="NY - deaths - ARIMA")

STATES_OI %>%
  filter(state=="VT")%>%
  model(ETS(deaths)) %>%
  gg_tsresiduals()+
    labs(title="VT - deaths - ETS")
STATES_OI %>%
  filter(state=="VT")%>%
  model(ARIMA(deaths)) %>%
  gg_tsresiduals()+
  labs(title="VT - deaths - ARIMA")
```

